%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{section:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

As part of the \href{https://wiki.neic.no/wiki/Tryggve}{Tryggve
  project}, we presented a method to extend the local computational
resources, when they are all busy, with cloud resources of partners in
other countries.
%
This is an interesting alternative to buying more hardware, because we
stress-tested the network and disc access and we noticed that the
performance degradation is close to neglectable.
%
We show that the use of dedicated network together with a secure cloud
infrastructure allows running analyses on sensitive data (such as
genomic or health data).
%
We ran realistic pipelines on non-sensitive --but real-- data
successfully using compute resources that stretch across national
borders.
%

\paragraph{Future Work.}
%
It would be interesting to scale up the solution to \emph{many} nodes
in ePouta and \emph{some} nodes in Knox, to see
%how much the link can be shared.
at which point the communication link gets saturated (and vice-versa).
%
The network link is therefore a key component of the current
architecture.
%
Scaling up that aspect need not be linear. We see such a setup as
palliating a shortage of resources at a compute center, on either end
of the link. These situations should not happen often, so we feel it
is acceptable to have some small latency, rather than waiting for
local resources to be freed for a user's computations to take place.

%
Finally, regarding the implementation of the setup,
%
it is probably possible to tweak the NFS settings, or the TCP settings
in the kernel, to gain even further performance.
%
Moreover, it is worth updating altogether the Slurm and NFS solutions
with more recent tools, that could improve job scheduling and disk
access, such as (i)~the use object storage or (ii)~the use a separate
disk volume attached to the storage VM, rather than an ephemeral disk
on file (\ie the default in \codeline{libvirt}).
%
On the network side, broadcast traffic is still forwarded to all
interfaces on VLAN 1203 and therefore to all VMs. An improvment would
be to use OpenVSwitch to learn about MAC addresses and skip physical
nodes that do not host any VM for the given project. That will improve
\emph{east-west} traffic. Another alternative is to distribute the
router using DVR, which is however not available when using the Linux
Bridge mechanism.

